### 2018-05-31 dissemination.publication/08da9685582866a0/paper_abstract.tex

We present a customizable Collective Knowledge workflow with
an integrated scoreboard to automate multi-objective benchmarking,
exploration and co-design of diverse machine learning models,
frameworks, libraries and platforms.
%
We demonstrate how it can help to automatically find 
the most efficient stack for image classification 
on a Pareto frontier of accuracy, speed and size 
across MobileNets, TensorFlow, ArmCL OpenCL libraries,
and Arm Mali GPUs.
%
All artifacts, workflows and results from this paper
are shared as open-source, reproducible and reusable components
with a common JSON API for further validation, comparison
and improvement by the community.

These customizable and portable workflows can help enhance 
existing benchmarks such as MLPerf with plug\&play components 
and a live dashboard for a collaborative, reproducible and 
fair comparison of novel techniques, frameworks and platforms.
%
Such approach can help to guide further software and hardware improvements, 
automate Artifact Evaluation at conferences such as SysML, 
and accelerate technology transfer to industry.