### 2018-06-12 dissemination.publication/08da9685582866a0/paper_abstract.tex
Update abstract.

We present a customizable Collective Knowledge workflow to study the execution
time vs.\ accuracy trade-offs for the MobileNets CNN family.
%
We use this workflow to evaluate MobileNets on Arm Cortex CPUs using TensorFlow
and Arm Mali GPUs using several versions of the Arm Compute Library.
%
Our optimizations for the Arm Bifrost GPU architecture reduce the execution
time by 2--3 times, while lying on a Pareto-optimal frontier.
%
We also highlight the challenge of maintaining the accuracy when deploying CNN
models across diverse platforms.
%
We make all the workflow components (models, programs, scripts, etc.) publicly
available to encourage further exploration by the community.